The Cost of Undisturbed Landscapes

### 2020-05-13 projects/asparagus/doc/paper_clean.tex
uodate of project asparagus

    \section{Introduction} \label{sec:introduction}
    Future, low carbon energy systems are likely to largely rely on electricity from renewable energy technologies
    (RET), such as solar photovoltaics (PV) or wind turbines, according to a substantial body of
    literature~\citep[e.g.][]{Becker2014,Delucchi2011,Fernandes2014,Gils2017,Jacobson2011,Schmidt2016}.
    \todo{references not good for statement}
    A broad stream of papers~\citep[e.g.][]{Brown2018,Rodriguez2015,Schlachtberger2018,Scholz2017} investigates
    conditions under which total system cost, conceived as the sum of fuel and emission cost, operation and maintenance
    cost, and investment cost, can be minimized.
    In this context, available resource potentials are an important boundary condition for the determination of such
    least-cost system configurations.
    Fortunately, such resource potentials are assessed by numerous scholars at various levels.
    Frequently, these assessments combine indicators of the wind resource with technical or legal restrictions on land
    available for RET deployment~\citep[e.g.][]{Bosch2017,McKenna2015,Ryberg2020}.
    However, there is an increasing awareness that such approaches are not a proper reflection of deployable RET
    capacites, as a more nuanced analysis would have to account for the social acceptance of RET.
    Opposition to wind turbines, for example, is often based on perceived negative interference with
    landscapes~\cite{Mattmann2016}, potential harm to wildlife~\cite{Loss2013, Voigt2015, Wang2015}, or
    noise~\cite{Wang2015a}.
    Zerrahn~\cite{Zerrahn2017} provides an excellent survey of the literature on wind power externalities.

>>>>

    \section{Introduction} \label{sec:introduction}
    Limiting global warming to `well below 2^{\circ}C compared to pre-industrial levels', as agreed on in the `Paris
    agreement' requires rapid and substantial reductions in greenhouse gas emissions \citep{IPCC2019}.
    The European Union has committed itself to xxx.
    Within this framework, by 2030 Austria aims to generate 100\% of its electricity consumption from domestic renewable
    sources on annual balance\footnote{Industry own use and ancillary service provision is exempt.
    Together, these accounted for approximately 10\% of consumption in 2016.}.
    According to government estimates, additional electricity generation of \SI{27}{\tera\watt\hour} from renewable
    sources is required to meet policy goals.
    Out of this total increase in renewable electricity generation, \SI{1}{\tera\watt\hour} is thought to be sourced
    from biomass, and \SI{5}{\tera\watt\hour} from new hydropower generation, according to the government's climate and
    energy strategy \citep{BMNT2018}.
    The comparatively small projected increase in electricity generation from these sources results from the perception
    of biomass being ecologically \citep{Erb2018} and economically unsustainable, while Austria's vast hydropower
    potentials were already largely exploited \citep{Walder2014} even before the projected increase in hydro generation
    by \SI{5}{\tera\watt\hour}.
    As Austrians banned the use of nuclear power in a popular referendum in 1978 \citep{Pelinka1983}, wind and solar
    power are the only remaining options for a large-scale expansion of renewable power generation.

%    Erneuerbare sind notwendig
%    -> Wind und PV die einzigen Technologien mit großem Potential
%    -> Wind hat soziale Akzeptanzprobleme (bzw. negative welfare impacts)
%    -> Wind kann aber möglicherweise auch positive Effekte auf die Systemkosten haben
%    -> Keine Windkraft auszubauen und damit die Landschaft zu schützen hat also Opportunitätskosten
%    -> Wie hoch sind diese Opportunitätskosten?
%    -> Das ist interessant, weil man damit auch ausrechnen kann, wieviel man im Prinzip kompensieren könnte.
%    Fertig.

    Future, low carbon energy systems are likely to largely rely on electricity from renewable energy technologies
    (RET), such as solar photovoltaics (PV) or wind turbines, according to a substantial body of
    literature~\citep[e.g.][]{Becker2014,Delucchi2011,Fernandes2014,Gils2017,Jacobson2011,Schmidt2016}.
    \todo{references not good for statement}
    A broad stream of papers~\citep[e.g.][]{Brown2018,Rodriguez2015,Schlachtberger2018,Scholz2017} investigates
    conditions under which total system cost, conceived as the sum of fuel and emission cost, operation and maintenance
    cost, and investment cost, can be minimized.
    In this context, available resource potentials are an important boundary condition for the determination of such
    least-cost system configurations.
    Fortunately, such resource potentials are assessed by numerous scholars at various levels.
    Frequently, these assessments combine indicators of the wind resource with technical or legal restrictions on land
    available for RET deployment~\citep[e.g.][]{Bosch2017,McKenna2015,Ryberg2020}.
    However, there is an increasing awareness that such approaches are not a proper reflection of deployable RET
    capacites, as a more nuanced analysis would have to account for the social acceptance of RET.
    Opposition to wind turbines, for example, is often based on perceived negative interference with
    landscapes~\cite{Mattmann2016}, potential harm to wildlife~\cite{Loss2013, Voigt2015, Wang2015}, or
    noise~\cite{Wang2015a}.
    Zerrahn~\cite{Zerrahn2017} provides an excellent survey of the literature on wind power externalities.
